Relative parabolic factorizations of Chevalley groups

### 2015-09-05 bass-kolster.tex
Edits from 04.09.15 meeting
Iterated BK: the statement formatted into a table

The estimate in the case $\Phi=\rB_\ell$ can be obtained in a similar way. \end{proof}
The Bass---Kolster decomposition can be iterated if the stability assumption allows, with a slight improvement of Corollary~\ref{cor:bass-kolster-count}.
\begin{cor}\label{cor:bass-kolster-iterated}
Assume $\sr(I)\leqslant2$ in case $\Phi=\rA_\ell$, $\sr(I)\leqslant3$ in case $\rC_\ell$ and $\asr(I)\leqslant2$ in case $\Phi=\rB_\ell,\rD_\ell$. Then every element of $\G(\Phi,R,I)$ is a product of an element of $\G(\langle\pm\alpha_\ell\rangle,R,I)$, at most $\ell-1$ elements of the form $z_{-\alpha}(*,*)$, $\alpha$ the maximal root of some terminal subsystem containing $\alpha_\ell$, and at most $N$ elementary root unipotents $x_\alpha(s)$, $s\in I$, where $N$ is $3\left|\Phi^+\right|+\ell$ in case $\Phi=\rA_\ell$, $N=3\left|\Phi^+\right|+2\ell$ in case $\Phi=\rC_\ell$ and $N=4\left|\Phi^+\right|-\ell$ in case $\Phi=\rB_\ell,\rD_\ell$.
\end{cor}
\begin{proof}
The improved estimate for $\Phi=\rA_\ell$ or $\rC_\ell$ is obtained by using the fact that it suffices to make only two or three additions respectively to make the shorter column unimodular in the first step of the Bass---Kolster decomposition.
\end{proof}
\[
\begin{tikzcd}[row sep=tiny]
\rD_4 \arrow[out=0,in=120]{rd}{\asr\leqslant3} & & & \\
\rA_4 \arrow{r}{\sr\leqslant4} & \rA_3 \arrow{r}{\sr\leqslant3} & \rA_2 \arrow{r}{\sr\leqslant2} & \rA_1 \\
\rC_4 \arrow{r}{\sr\leqslant7} & \rC_3 \arrow{r}{\sr\leqslant5} & \rC_2 \arrow[out=0,in=-120,swap]{ru}{\sr\leqslant3} \\
\rB_4 \arrow{r}{\asr\leqslant3} & \rB_3 \arrow[out=0,in=-120,swap]{ru}{\asr\leqslant2} & &
\end{tikzcd}
\]
\begin{proof}[Proof of Theorem~\ref{thm:SL2width}]
A direct application of Corollary~\ref{cor:bass-kolster-iterated}.
\end{proof}

>>>>

The estimate in the case $\Phi=\rB_\ell$ can be obtained in a similar way. \end{proof}
The Bass---Kolster decomposition can be iterated if the stability assumption allows, with a slight improvement of Corollary~\ref{cor:bass-kolster-count}.
\begin{cor}\label{cor:bass-kolster-iterated}
Let $\Phi$ be a root system, $R$ a commutative ring and $I$ its ideal. Then under a suitable stability assumption every element of $\G(\Phi,R,I)$ is a product of an element of $\G(\langle\pm\alpha_\ell\rangle,R,I)\cong\SL(2,R,I)$, at most $\ell-1$ elements of the form $z_{-\alpha}(*,*)$, $\alpha$ the maximal root of some terminal subsystem containing $\alpha_\ell$, and at most $N$ elementary root unipotents $x_\alpha(s)$, $s\in I$. The stability assumption and the number $N$ are as follows:
\[\begin{array}{l@{\quad}l@{\quad}l}
\Phi=\rA_\ell, & \sr(I)\leqslant 2, & N=3\left|\Phi^+\right|+\ell; \\
\Phi=\rC_\ell, & \sr(I)\leqslant 3, & N=3\left|\Phi^+\right|+2\ell; \\
\Phi=\rB_\ell,\rD_\ell, & \asr(I)\leqslant 2, & N=4\left|\Phi^+\right|-\ell.
\end{array}\]
\end{cor}
\begin{proof}
The improved estimate for $\Phi=\rA_\ell$ or $\rC_\ell$ is obtained by using the fact that it suffices to make only two or three additions respectively to make the shorter column unimodular in the first step of the Bass---Kolster decomposition.
\end{proof}
\begin{figure}[htb]\label{fig:bass-kolster}
\[
\begin{tikzcd}[row sep=tiny]
\rD_4 \arrow[out=0,in=120]{rd}{\asr\leqslant3} & & & \\
\rA_4 \arrow{r}{\sr\leqslant4} & \rA_3 \arrow{r}{\sr\leqslant3} & \rA_2 \arrow{r}{\sr\leqslant2} & \rA_1 \\
\rC_4 \arrow{r}{\sr\leqslant7} & \rC_3 \arrow{r}{\sr\leqslant5} & \rC_2 \arrow[out=0,in=-120,swap]{ru}{\sr\leqslant3} \\
\rB_4 \arrow{r}{\asr\leqslant3} & \rB_3 \arrow[out=0,in=-120,swap]{ru}{\asr\leqslant2} & &
\end{tikzcd}
\]
\caption{Reductions provided by the Bass---Kolster decomposition}
\end{figure}

\begin{proof}[Proof of Theorem~\ref{thm:SL2width}]
A direct application of Corollary~\ref{cor:bass-kolster-iterated}.
\end{proof}
